Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections at the LArIAT experiment

### 2018-06-25 AppendixB-EField/B-LArIATEField.tex
s

Before getting into the details of the measurement procedures, it is important to explicit the relationship between some  quantities in play. The electric field and the drift velocity ($v_{drift}$) are related as follows 
\begin{equation} v_{drift} = \mu(E_{field},T) E_{field}, \label{eq:vd}
\end{equation}
where $\mu$ is the electron mobility, which depends on the electric field and on the temperature (T). The empirical formula for this dependency is described in ~\cite{WWW} and shown in Figure \ref{fig:EV} for several argon temperatures.

@@

The relationship between the drift time ($t_{drift}$) and the drift velocity is trivially given by
\begin{equation}
t_{drift} = \Delta x/v_{drift}, \label{eq:drifttime}
\end{equation}
where $\Delta x$ is the distance between the edges of the drift region.
Table \ref{tab:Efields} reports the values of the electric field, drift velocity, and drift times for the smaller drift volumes.

@@

\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/CircuitLArIAT.png}
\caption{LArIAT HV simple schematics.}
\label{fig:circuit}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/glassman_current_20160525-30.png}
\caption{Current reading from the Glassman between May 25th and May 30th, 2016 (typical Run-II conditions).}
\label{fig:currentMeasurement}
\end{minipage}
\end{figure}

@@

\subsection*{E field using cathode-anode piercing tracks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[b]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/TPCCrossSectionView.png}
\caption{Pictorial representation of the YX view of the TPC. The distance within the anode planes and between the shield plane and the cathode is purposely out of proportion to illustrate the time difference between hits on collection and induction. An ACP track is shown as an example.}
\label{fig:Scheme}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/AngleDef.png}
\caption{Angle definition in the context of LArIAT coordinate system.}
\label{fig:AngleDef}
\end{minipage}
\end{figure}
We devise an independent method to measure the drift time (and consequently drift velocity and electric field) using TPC cathode to anode piercing tracks. We use this method as a cross check to the single line method.
The basic idea is simple:
\begin{itemize}
\item[0.] Select cosmic ray events with only 1 reconstructed track 
\item[1.] Reduce the events to the one containing tracks that cross both anode and cathode
\item[2.] Identify the first and last hit of the track
\item[3.] Measure the time difference between these two hits ($\Delta t$).
\end{itemize}
This method works under the assumptions that the time it takes for a cosmic particle to cross the chamber ($\sim$ns) is small compared to the charge drift time ($\sim$ hundreds of $\mu$s).

We choose cosmic events to allow for a high number of anode to cathode piercing tracks (ACP tracks), rejecting  beam  events where the particles travel almost perpendicularly to drift direction. We select events with only one reconstructed track to  maximize the chance of selecting a single crossing muon (no-michel electron). We utilize ACP tracks because their hits span the full drift length of the TPC, see figure \ref{fig:Scheme}, allowing us to define where the first and last hit of the tracks are located in space regardless of our assumption of the electric field. %The definition of the last hit is easy: it is the hit closest to the cathode. The definition of the first hit is a bit more complicated. A track that crosses the anode planes will deposit charge in the small drift volumes (S-I and I-C). The drift time in S-I and I-C region was already calculated in Table \ref{tab:Efields}. This is to say that the position of the first hit matters when calculating the drift time at the microsecond precision. Single hits on the collection plan will not form a 3D object. This means that we can safely exclude that the reconstruction of ACP tracks starts at the cathode plane (black point in figure \ref{fig:Scheme}). %Understanding if the first hit of a track in on the induction or on the shield plan is more complicated. For now, we'll take an uncertainty hit of 2.3 $\mu$s.

@@

For each track passing our selection, we loop through the associated hits to retrieve the timing information. The analysis is performed separately on hits on the collection plane and induction plane, but lead to consistent results. As an example of the time difference, figures \ref{fig:Run2PosColFit} and \ref{fig:Run2PosIndFit} represent the difference in time between the last and first hit of the selected tracks for Run-II Positive Polarity sample on the collection and induction plane respectively.  We fit with a Gaussian to the peak of the $\Delta t$ distributions to extract the mean drift time and the uncertainty associated with it. The long tail at low $\Delta t$ represents contamination of non-ACP tracks in the track selection.  We apply the same procedure to Run-I and Run-II, positive and negative polarity alike.

>>>>

Before getting into the details of the measurement procedures, it is important to clarify the relationship between the  quantities in play. The electric field ($E_{field}$) and the drift velocity ($v_{drift}$) are related as follows 
\begin{equation} v_{drift} = \mu(E_{field},T) E_{field}, \label{eq:vd}
\end{equation}
where $\mu$ is the electron mobility, which depends on the electric field and on the temperature (T). The empirical formula for this dependency is described in ~\cite{WWW} and shown in Figure \ref{fig:EV} for several argon temperatures.

@@

The relationship between the drift time ($t_{drift}$) and the drift velocity is trivially given by
\begin{equation}
t_{drift} = \Delta x/v_{drift}, \label{eq:drifttime}
\end{equation}
where $\Delta x$ is the distance between the edges of the drift region.
Table \ref{tab:Efields} reports the values of the electric field, drift velocity, and drift times for the TPC's small drift volumes.

@@

\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/CircuitLArIAT.png}
\caption{LArIAT HV simple scheme.}
\label{fig:circuit}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/glassman_current_20160525-30.png}
\caption{Current reading from the Glassman between May 25th and May 30th, 2016 (typical Run-II conditions).}
\label{fig:currentMeasurement}
\end{minipage}
\end{figure}

@@

\subsection*{E field using cathode-anode piercing tracks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[b]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/TPCCrossSectionView.png}
\caption{Pictorial representation of the YX view of the TPC. The distance within the anode planes and between the shield plane and the cathode is purposely out of proportion to illustrate the time difference between hits on collection and induction. An ACP track is shown as an example.}
\label{fig:Scheme}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=3in]{AppendixB-EField/Images/AngleDef.png}
\caption{Angle definition in the context of LArIAT coordinate system.}
\label{fig:AngleDef}
\end{minipage}
\end{figure}
We devised an independent method to measure the drift time (and consequently drift velocity and electric field) using TPC cathode to anode piercing tracks. We use this method as a cross check to the single line method.
The basic idea is simple:
\begin{itemize}
\item[1.] Select cosmic ray events with only 1 reconstructed track 
\item[2.] Reduce the events to the ones containing tracks that cross both anode and cathode
\item[3.] Identify the first and last hit of the track
\item[4.] Measure the time difference between these two hits ($\Delta t$).
\end{itemize}
This method works under the assumptions that the time it takes for a cosmic particle to cross the chamber ($\sim$ns) is small compared to the charge drift time ($\sim$ hundreds of $\mu$s).

We choose cosmic events to allow for a high number of anode to cathode piercing tracks (ACP tracks), rejecting  beam  events where the particles travel almost perpendicularly to drift direction. We select events with only one reconstructed track to  maximize the chance of selecting a single crossing muon (no michel electron). We utilize ACP tracks because their hits span the full drift length of the TPC, see Figure \ref{fig:Scheme}, allowing us to define where the first and last hit of the tracks are located in space regardless of our assumption of the electric field. %The definition of the last hit is easy: it is the hit closest to the cathode. The definition of the first hit is a bit more complicated. A track that crosses the anode planes will deposit charge in the small drift volumes (S-I and I-C). The drift time in S-I and I-C region was already calculated in Table \ref{tab:Efields}. This is to say that the position of the first hit matters when calculating the drift time at the microsecond precision. Single hits on the collection plan will not form a 3D object. This means that we can safely exclude that the reconstruction of ACP tracks starts at the cathode plane (black point in figure \ref{fig:Scheme}). %Understanding if the first hit of a track in on the induction or on the shield plan is more complicated. For now, we'll take an uncertainty hit of 2.3 $\mu$s.

@@

For each track passing our selection, we loop through the associated hits to retrieve the timing information. The analysis is performed separately on hits on the collection plane and induction plane, but leads to consistent results. As an example of the time difference, figures \ref{fig:Run2PosColFit} and \ref{fig:Run2PosIndFit} represent the difference in time between the last and first hit of the selected tracks for Run-II Positive Polarity sample on the collection and induction plane respectively.  We fit with a Gaussian to the peak of the $\Delta t$ distributions to extract the mean drift time and the uncertainty associated with it. The long tail at low $\Delta t$ represents contamination of non-ACP tracks in the track selection.  We apply the same procedure to Run-I and Run-II, positive and negative polarity alike.

### 2018-06-25 AppendixB-TrackingStudies/B-Tracking.tex
s

Scope of this study is assessing the performances of the WC2TPC match on Monte Carlo (see Section \ref{ch:WC2TPCMatchMethod} ) and decide the selection values we will use on data. A word of caution is necessary here. With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g. the incorrect association between the beamline hadron and its decay products inside the TPC.  Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.

@@

We start by defining a figure of merit in order to discern what makes a parameter configuration better than an other. We chose the percentage of events whose reconstructed and true length differ less than 2 cm. We then identify the parameters in TrajCluster that are most important to  correctly stop  the tracking and an appropriate range of values for each of them. We chose to optimize the parameters that leverage on the angle between consecutive groups of hits, the number of hits use in the cluster fit and the average hit charge to stop the tracking. We define a configuration space with all  possible combination of values for the chosen parameters and we perform reconstruction one combination at a time: the combination with the highest figure of merit determines the optimized tracking reconstruction.

We chose construct the combination space using a total of 5 parameters, 3 values each and two iterations of the method (for a total of 486 combinations). We run the combinations on a sample of 100000 pion events. 
After the optimization, the most upstream point of the tracking is correctly identified 99.5\% of the times, the most downstream point is correctly identified 62.5\% of the times, the tracking stops short about 15\% of the times and misses the interaction point 22.5\% of the times. Hand scanning confirmed that the missed interaction points happen in the vast majority of cases for very shallow angles, as shown in the event display in Figure, or in the case of angles visible only in one projection plane. We also noticed that the premature stopping of the tracks is often related to the presence of delta rays parallel to the track. We see room of improvement, such as the delta ray removal and a forced track breaking in case of a kink present in a single plane, for a future analysis. \textcolor{red}{ADD evd}

@@

>>>>

Scope of this study is assessing the performances of the WC2TPC match on Monte Carlo (see Section \ref{ch:WC2TPCMatchMethod}) and decide the selection values we will use on data. A word of caution is necessary here. With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g. the incorrect association between the beamline hadron and its decay products inside the TPC.  Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.

@@

We start by defining a figure of merit in order to discern what makes a parameter configuration better than an other. We chose the percentage of events whose reconstructed and true length differ less than 2 cm. In order to identify the true interaction point, no selection is performed on the scattering angle: even hadronic interaction flagged by Geant4 with very shallow angles (< 5$^\circ$) are kept to perform the optimization.
We then identify the parameters in TrajCluster that are most important to  correctly stop  the tracking and an appropriate range of values for each of them. We chose to optimize the parameters that leverage on the angle between consecutive groups of hits, the number of hits use in the cluster fit and the average hit charge to stop the tracking. We define a configuration space with all  possible combination of values for the chosen parameters and we perform reconstruction one combination at a time: the combination with the highest figure of merit determines the optimized tracking reconstruction.

We chose construct the combination space using a total of 5 parameters, 3 values each and two iterations of the method (for a total of 486 combinations). We run the combinations on a sample of 100000 pion events. 
After the optimization, the most upstream point of the tracking is correctly identified 99.5\% of the times, the most downstream point is correctly identified 62.5\% of the times, the tracking stops short about 15\% of the times and misses the interaction point 22.5\% of the times. Hand scanning confirmed that the missed interaction points happen in the vast majority of cases for very shallow angles, as shown in the event display in Figure \ref{evd:kink}, or in the case of angles visible only in one projection plane. We also noticed that the premature stopping of the tracks is often related to the presence of delta rays parallel to the track. We see room of improvement, such as the delta ray removal and a forced track breaking in case of a kink present in a single plane, for a future analysis.

@@

\begin{figure}[hpbt]
\centering
\includegraphics[width=15cm]{AppendixB-TrackingStudies/Kink.png}
\caption{Example of a shallow angle hadronic interaction ``missed" by the TrajCluster.}
\label{evd:kink}
\end{figure}

### 2018-06-25 AppendixC-EnergyCalibration/C-EnergyCalibration.tex
s

Scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to energy deposited in the chamber (dE). As described in section \ref{sec:SignalProc}, this is a multi-step procedure. In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{LArIATdqdx}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values. Lastly, we apply overall calibration of the energy, i.e. we determine the ``calorimetry constants" using the procedure described in this section.

@@

The basic idea of this calibration technique is to utilize a sample of beamline events with known particle species and momentum to measure the $dE/dX$ of the corresponding tracks in the TPC. In particular, we decided to use positive pions as calibration sample and samples from all the other particle species as cross check. Once the $dE/dX$ of the positive pion sample  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}.

In data, we start by selecting a sample of beamline positive pion beamline candidates without any restriction on their measured momentum\footnote{it should be noted that some muon and position contamination is present in the $\pi^+$ sample}.
We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face. For each surviving pion candidate,  we measure the $dE/dx$ at each of the first 12 spacepoints associated the 3D reconstructed track, corresponding to a $\sim$ 5 cm portion. These $dE/dX$ measurements are then put into a histogram that corresponds to measured momentum of the track. The $dE/dX$ histograms are sampled every 50 MeV/c in momentum (e.g. 150~MeV/c $< P <$ 200~MeV/c, 200~MeV/c $< P <$ 250/c~MeV, etc...).   This process of selecting, sampling, and recording the $dE/dX$ for various momentum bins is repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV/c and 1100~MeV/c. On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV. Thus choosing 50 MeV/c size bins for our histograms covers the energy spread within those bins due to energy loss from ionization for all the particle species identifiable in the beamline. 
Each 50 MeV/c momentum binned $dE/dX$ histogram is now fit with a simple Landau function. The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted against the theoretical prediction Figure \ref{fig:PDGEnergyLossArgon}. Depending on the outcome of the data-prediction comparison, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.  We perform this  tuning for the collection and induction plane separately. 
As a cross check to the calorimetry constants determined using the positive pions, we lock the constants and  plot the $dE/dx$ versus momentum distribution of all the other particle species identifiable in the beamline data ($\pi/\mu/e$, K , p, in both polarities) against the corresponding Beth-Bloch prediction. The agreement between data from the other particle species and the predictions is the expected result of this cross check.
The results of the tuning and cross check for Run-II data on the collection plane is shown in Figure \ref{fig:BBandData}  negative polarity data on top, positive polarity data on the bottom.

@@

After the calibration is done separately on data and MC, we can compare the resulting dE/dX distributions; this is done for the set of pion beamline candidates and pion MC on the left side of Figure \ref{fig:dedx}. On the right side of the same figure, we report the data-MC comparison for the kaon sample used in the cross section. The distributions are fitted with simple Landau functions. As expected, the Landau MPV for data  is consistent with the MC MPV for both pions and kaons. For both the pion and the kaon case, the width of the Landau is a bit wider in data than in MC; this difference might be due to an underestimate of the electronic noise in the MC simulation.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{AppendixC-EnergyCalibration/dEdXPions.pdf}
\includegraphics[width=0.48\textwidth]{AppendixC-EnergyCalibration/dEdXKaons.pdf}
\caption{\emph{Left:} dE/dx distribution for $\pi^-/\mu^-/e^-$ data (black) and Pion MC (blue). The Landau fit for data is shown in red, the one for MC in teal.\emph{Right:} dE/dx distribution for $K^+$ data (black) and Kaon MC (blue). The Landau fit for data is shown in red, the one for MC in teal. All the distributions are area normalized.}
\label{fig:dedx}
\end{figure}

>>>>

The scope of the energy calibration is to identify the factors which convert the charge collected (dQ) to the energy deposited in the chamber (dE). As described in Section \ref{sec:SignalProc}, this is a multi-step procedure. In LArIAT, we first correct the raw charge by the electronic noise on the considered wire \cite{LArIATdqdx}, then by the electron lifetime \cite{LArIATLifeTime},  and then by the recombination using the ArgoNeut recombination values. Lastly, we apply overall calibration of the energy, i.e., we determine the ``calorimetry constants" using the procedure described in this section.

@@

The basic idea of this calibration technique is to utilize a sample of beamline events with known particle species and momentum to measure the $dE/dX$ of the corresponding tracks in the TPC. In particular, we decided to use positive pions as the calibration sample and samples from all the other particle species as a cross check. Once the $dE/dX$ of the positive pion sample  has been measured at various momenta, we tune to calorimetry constants within the reconstruction software to align the measured values to match the theoretical ones found in Figure \ref{fig:PDGEnergyLossArgon}. This is a new technique for calorimetry calibration in liquid argon developed by a joint ArgoNeut and  LArIAT effort.

In data, we start by selecting a sample of beamline positive pion beamline candidates without any restriction on their measured momentum\footnote{It should be noted that some muon and position contamination is present in the $\pi^+$ sample.}.
We then apply the WC2TPC match and subtract the energy loss upstream to the TPC front face, determining the momentum at the TPC front face. For each surviving pion candidate,  we measure the $dE/dx$ at each of the first 12 spacepoints associated the 3D reconstructed track, corresponding to a $\sim$ 5 cm portion. These $dE/dX$ measurements are then put into a histogram that corresponds to measured momentum of the track. The $dE/dX$ histograms are sampled every 50 MeV/c in momentum (e.g. 150~MeV/c $< P <$ 200~MeV/c, 200~MeV/c $< P <$ 250/c~MeV, etc...).   This process of selecting, sampling, and recording the $dE/dX$ for various momentum bins is repeated over the entire sample of events, allowing us to collect sufficient statistic in most of the momentum bins between 150~MeV/c and 1100~MeV/c. On average, pions and muons only lose $\sim$10 MeV in this 5~cm section of the track and protons lose $\sim$20 MeV. Thus choosing 50 MeV/c size bins for our histograms covers the energy spread within those bins due to energy loss from ionization for all the particle species identifiable in the beamline. 
Each 50 MeV/c momentum binned $dE/dX$ histogram is now fit with a simple Landau function. The most probable value (MPV) and the associated error on the MPV from the fit are extracted and plotted against the theoretical prediction Figure \ref{fig:PDGEnergyLossArgon}. Depending on the outcome of the data-prediction comparison, we modify the calorimetry constants and we repeat the procedure until a qualitative agreement is achieved.  We perform this  tuning for the collection and induction plane separately. 
As a cross check to the calorimetry constants determined using the positive pions, we lock the constants and  plot the $dE/dx$ versus momentum distribution of all the other particle species identifiable in the beamline data ($\pi/\mu/e$, K , p, in both polarities) against the corresponding Bethe-Bloch prediction. The agreement between data from the other particle species and the predictions is the expected result of this cross check.
The results of the tuning and cross check for Run-II data on the collection plane is shown in Figure \ref{fig:BBandData}  negative polarity data on top, positive polarity data on the bottom.

@@

After the calibration is done separately on data and MC, we can compare the resulting dE/dX distributions; this is done for the set of pion beamline candidates and pion MC used in the cross section analysis on the left side of Figure \ref{fig:dedx}. On the right side of the same figure, we report the data-MC comparison for the kaon sample used in the cross section analysis. The distributions are fitted with simple Landau functions. As expected, the Landau MPV for data  is consistent with the MC MPV for both pions and kaons. For both the pion and the kaon case, the width of the Landau is a bit wider in data than in MC; this difference might be due to an underestimate of the electronic noise in the MC simulation.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{AppendixC-EnergyCalibration/dEdXPions.pdf}
\includegraphics[width=0.48\textwidth]{AppendixC-EnergyCalibration/dEdXKaons.pdf}
\caption{\emph{Left:} dE/dx distribution for $\pi^-/\mu^-/e^-$ data (black) and Pion MC (blue). The Landau fit for data is shown in red, the one for MC in teal. \emph{Right:} dE/dx distribution for $K^+$ data (black) and Kaon MC (blue). The Landau fit for data is shown in red, the one for MC in teal. All the distributions are area normalized.}
\label{fig:dedx}
\end{figure}

### 2018-06-25 Chapter-0/0-Introduction.tex
s

Experimentally, precision measurements can be achieved only if the detector technology is able to resolve the fine details of a neutrino interaction and to record a statistically relevant number of neutrinos.  With ``fine details" here we mean the ability to distinguish the many products of the neutrino interaction, such as protons, pions, muons and electrons, and to measure their energy.
Historically,  bubble chamber neutrino detectors were the first revolution in neutrino detection: for example, the spatial resolution of Gargamelle allowed the discovery of neutrino neutral current interaction\cite{HASERT1973138}. Despite the high precision of bubble chambers images, this technology is hard to scale to massive size, making statistical analyses on neutrino interactions almost impossible to perform. To make up for the small neutrino interaction cross section, neutrino experiments moved to very large size, at the expenses of spatial precision. This is the case for the detectors which discovered neutrino oscillation:  both Super-Kamiokande and SNO are massive Cherenkov detectors. With LArTPCs, the field is gaining again bubble-chamber like precision but at massive scales. Following the recommendations of  the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the intensity frontier through the construction of massive LArTPCs. In particular, we are seeing the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program  (DUNE), both based on the LArTPC detector technology. The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino? is CP violated in the lepton sector? are there any additional symmetries? and, can we find an indication of Grand Unified Theories?

@@

DUNE has a vast neutrino and non-accelerator physics reach. For what it concerns neutrino physics, oscillation analyses in DUNE have the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector. Besides its neutrino program, DUNE can open an experimental window on Grand Unified Theories (GUTs). GUTs could potentially answer fundamental questions such as the existence of non-zero neutrino masses and matter-antimatter asymmetry, explaining some ``accidents" in the Standard Models, such as the exact cancellation of the  proton and the electron charge.   Directly probing GUTs at the unification energy scale is impossible by any foreseeable collider experiment. We then need an indirect proof such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales. Historically, the main technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years. The DUNE far detector and its non-accelerator physics program is a interesting new actor on this stage.  LArTPCs can in fact complement nucleon decay searches in modes where water Cherenkov detectors are less sensitive, especially $p\rightarrow K^+\bar{\nu}$  \cite{Adams:2013qkq}.

@@

The amount of information a LArTPC can provide makes these detectors rather complex: a series of dedicated measurements is necessary to obtain meaningful physics results from a LArTPC. The complexity of the LArTPC technology for neutrino detection is due to several reasons. Argon is a fairly heavy element, which means that nuclear effects play an important role in the interaction topology. For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of pion in argon. %No measurement of hadronic cross sections for pions or kaons is available for argon (yet). 
The amount of details in a LArTPC event can easily be parsed by human eye, but can make automatic event reconstruction rather challenging. Thus, reconstruction algorithms in LArTPC need to be tuned to recognize the different topologies of the neutrino interaction products in argon. This is particularly true for pions, since they are copiously produced of the neutrino interactions: the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction channel.

>>>>

Experimentally, precision measurements can be achieved only if the detector technology is able to resolve the fine details of a neutrino interaction and to record a statistically relevant number of neutrinos.  With ``fine details" here we mean the ability to distinguish the many products of the neutrino interaction, such as protons, pions, muons and electrons, and to measure their energy.
Historically,  bubble chamber neutrino detectors were the first revolution in neutrino detection: for example, the spatial resolution of Gargamelle allowed the discovery of neutrino neutral current interaction\cite{HASERT1973138}. Despite the high precision of bubble chambers images, this technology is hard to scale to massive size, making statistical analyses on neutrino interactions almost impossible to perform. To make up for the small neutrino interaction cross section, neutrino experiments moved to very large size, at the expenses of spatial precision. This is the case for the detectors which discovered neutrino oscillation:  both Super-Kamiokande and SNO are massive Cherenkov detectors \cite{PDGOsc}. With LArTPCs, the field is gaining again bubble-chamber like precision but at massive scales. Following the recommendations of  the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US particle physics panorama is directing a substantial effort towards the exploration of the intensity frontier through the construction of massive LArTPCs. In particular, we are seeing the development of a Short Baseline Neutrino Program (SBN) and long baseline neutrino program  (DUNE), both based on the LArTPC detector technology. The US liquid argon program has the potential to answer many of the fundamental open questions in particle physics today, such as: is there a fourth generation neutrino? is CP violated in the lepton sector? are there any additional symmetries? and, can we find an indication of Grand Unified Theories?

@@

DUNE has a vast neutrino and non-accelerator physics reach. Within neutrino physics, oscillation analyses in DUNE have the capability of solving the mass hierarchy and octant problem,  and discovering CP violation in the neutrino sector. Besides its neutrino program, DUNE can open an experimental window on Grand Unified Theories (GUTs). GUTs could potentially answer fundamental questions such as the existence of non-zero neutrino masses and matter-antimatter asymmetry, explaining some ``accidents" in the Standard Models, such as the exact cancellation of the  proton and the electron charge.   Directly probing GUTs at the unification energy scale is impossible by any foreseeable collider experiment. We then need an indirect proof such as baryon number violation, which is predicted by almost every GUT in the form of proton decay, bounded nucleon decay or $n-\bar n$ oscillations on long time-scales. Historically, the main technology used in these searches has been water Cherenkov detectors, with Super-Kamiokande setting all the current experimental limits on the decay lifetimes at the order of $\sim 10 ^{34}$ years. The DUNE far detector and its non-accelerator physics program is a interesting new actor on this stage.  LArTPCs can in fact complement nucleon decay searches in modes where water Cherenkov detectors are less sensitive, especially $p\rightarrow K^+\bar{\nu}$  \cite{Adams:2013qkq}.

@@

The amount of information a LArTPC can provide makes these detectors rather complex: a series of dedicated measurements is necessary to obtain meaningful physics results from a LArTPC. The complexity of the LArTPC technology for neutrino detection is due to several reasons. Argon is a fairly heavy element, which means that nuclear effects play an important role in the interaction topology. For example, pions are one of the main products of neutrino interactions; yet,  since data on charged particle interaction in argon is scarce, neutrino event generators have big uncertainties in the re-scattering simulation of pions in argon. %No measurement of hadronic cross sections for pions or kaons is available for argon (yet). 
The amount of details in a LArTPC event can easily be parsed by human eye, but can make automatic event reconstruction rather challenging. Thus, reconstruction algorithms in LArTPC need to be tuned to recognize the different topologies of the neutrino interaction products in argon. This is particularly true for pions, since they are copiously produced of the neutrino interactions: the occurrence of a pion interaction in argon can modify the topology of the neutrino event, causing a misidentification of the neutrino interaction channel.

### 2018-06-25 Chapter-1/1-Theory.tex
s

\subsection{Neutrinos in the Standard Model}
Neutrino can be introduced in the SM as left-handed massless Weyl spinors.
The Dirac equation of motion for a free field
\begin{equation}
(i\gamma^ \mu \partial_\mu - m) \psi = 0
\end{equation}
for a fermionic field 
\begin{equation}
 \psi =  \psi_L +  \psi_R
\end{equation}
is equivalent to the equations
\begin{equation}
i\gamma^ \mu \partial_\mu  \psi_L = m \psi_R
\label{eq:15}
\end{equation}
\begin{equation}
i\gamma^ \mu \partial_\mu  \psi_R = m \psi_L
\label{eq:16}
\end{equation}

@@

where $c$ and $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$  are the eventual Majorana CP violation phases.  Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = m^2_{a} - m^2_{b}$, where $a$ and $b$ represent the mass eigenstates. A summary of the current status of experimental results is given in table \ref{tab:nuosc}.

\begin{table}[htpb]
\centering
\caption{Summary of experimental results on neutrino oscillation parameters. \textcolor{red}{ADD CITATIONS}}
\label{tab:nuosc}
\begin{tabular}{|r|c|c|c|}
\hline
        & Value   & Precision & Experiment \\
\hline
$\theta_{23}$            & $\sim$45$^\circ$                      & 9.0\%   & Super Kamiokande, MINOS,\\
|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  NO$\nu$a, MACRO, Super Kamiokande\\
\hline
$\theta_{12}$            & 34$^\circ$                     & 5.8\% &  SNO,Gallex,\\
|$\Delta m^2_{12}$|    & 7.4 $10^{-5}$ eV$^2$   & 2.8\% &  SAGE,  KamLAND \\
\hline
$\theta_{13}$            &   9$^\circ$                    & 4.7\%  & Daya  Bay, \\
|$\Delta m^2_{32}$|    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure}[hbpt]
\centering
\includegraphics[width=0.7\textwidth]{Chapter-1/Images/KaonsEvd.pdf}
\caption{Kaon-Argon interaction topologies due to elastic scattering and inelastic scattering as seen in LArIAT data (collection plane only).}
\label{fig:KaonsEvd}
\end{figure}

>>>>

\subsection{Neutrinos in the Standard Model}
Neutrinos can be introduced in the SM as left-handed massless Weyl spinors.
The Dirac equation of motion for a free field
\begin{equation}
(i\gamma^ \mu \partial_\mu - m) \psi = 0
\end{equation}
for a fermionic field 
\begin{equation}
 \psi =  \psi_L +  \psi_R
\end{equation}
is equivalent to the equations
\begin{equation}
i\gamma^ \mu \partial_\mu  \psi_L = m \psi_R
\label{eq:15}
\end{equation}
\begin{equation}
i\gamma^ \mu \partial_\mu  \psi_R = m \psi_L
\label{eq:16}
\end{equation}

@@

where $c$ and $s$ stand respectively for cosine and sine of the corresponding  mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), $\delta$ is the Dirac CP violation phase, $\alpha_1$ and $\alpha_{2}$  are the eventual Majorana CP violation phases.  Experimental results on neutrino oscillations are generally reported in terms of the mixing angles and of the squared mass splitting $\Delta m^2_{ab} = m^2_{a} - m^2_{b}$, where $a$ and $b$ represent the mass eigenstates. A summary of the current status of experimental results extracted from \cite{PDGOsc} is given in table \ref{tab:nuosc}.

\begin{table}[htpb]
\centering
\caption{Summary of experimental results on neutrino oscillation parameters extracted from \cite{PDGOsc}.}
\label{tab:nuosc}
\begin{tabular}{|r|c|c|c|}
\hline
        & Value   & Precision & Experiment \\
\hline
$\theta_{23}$            & $\sim$45$^\circ$                      & 9.0\%   & Super Kamiokande, MINOS,\\
$|\Delta m^2_{32}|$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  NO$\nu$a, MACRO, Super Kamiokande\\
\hline
$\theta_{12}$            & 34$^\circ$                     & 5.8\% &  SNO,Gallex,\\
$|\Delta m^2_{12}|$    & 7.4 $10^{-5}$ eV$^2$   & 2.8\% &  SAGE,  KamLAND \\
\hline
$\theta_{13}$            &   9$^\circ$                    & 4.7\%  & Daya  Bay, \\
$|\Delta m^2_{32}|$    & 2.5 $10^{-3}$ eV$^2$   & 1.8\% &  RENO\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure}[hbpt]
\centering
\includegraphics[width=\textwidth]{Chapter-1/Images/KaonsEvd.pdf}
\caption{Kaon-Argon interaction topologies due to elastic scattering and inelastic scattering as seen in LArIAT data (collection plane only).}
\label{fig:KaonsEvd}
\end{figure}
